---
title: "ADS506 Team 4 Project"
author: "Jacqueline Urenda"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r}
#loading Libraries:
library(tidyverse)
library(fpp2)
library(readr)
library(forecast)
library(ggplot2)
library(gridExtra)
library(reshape2)
library(dplyr)
library(lubridate)
set.seed(506)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Exploratory Data Analysis: 

```{r}
# Loading the Tetuan City Power Consumption Dataset

pc_df <- read_csv("/Users/jackieurenda/Desktop/ADS 506/Tetuan City power consumption.csv", 
                  col_types = cols(DateTime = col_datetime(format = "%m/%d/%Y %H:%M")))

head(pc_df)
tail(pc_df)
```

EDA and Understanding Data:
```{r}
sum(is.na(pc_df))
#there are no null values in the dataset
```
Descriptive Summary Statisics:
```{r}

summary(pc_df)

```
Dataset dates range from 01/01/2017 to 12/30/2017 (364 days total). The temperature median and mean are around 18 degrees celsius. Average humidity is 68.26%, which is pretty high. 



```{r}
library(corrplot)
corr <- round(x = cor(pc_df[, 2:9]), digits = 2)
library(RColorBrewer)
corrplot(corr, method = "color", outline = T, addgrid.col = "darkgray", order="hclust", addrect = 4, rect.col = "black", rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", tl.cex = .75, cl.cex = .75, addCoef.col = "white", number.digits = 2, number.cex = 0.5, col = colorRampPalette(c("darkred","white","midnightblue"))(100))
```
All three zones have strong correlations between each other. Temperature and wind speed have a correlation of 0.48. Humidity and temperature have a moderate but noticeable correlation as well. General diffuse flows and Diffuse flows have a correlation of 0.56.

Time Series Data on 10 Minute Interval: 
```{r}
#setting time series for each zone:
pc_ts1 <- ts(pc_df$`Zone 1 Power Consumption`)
pc_ts2 <- ts(pc_df$`Zone 2  Power Consumption`)
pc_ts3 <- ts(pc_df$`Zone 3  Power Consumption`)
```


PLotting Time Series Data individually:
```{r}
plot(pc_ts1, range = range(pc_ts1), col = 'blue', xlab = "Time", ylab ="Power Consumption")
plot(pc_ts2, range = range(pc_ts2), col = 'red', xlab = "Time", ylab ="Power Consumption")
plot(pc_ts3, range = range(pc_ts3), col = 'black', xlab = "Time", ylab ="Power Consumption")
```




Looking at distribution of columns:
```{r}
par(mfrow=c(3,3))
hist(pc_df$Temperature, main="Histogram of Temperature", col = "blue", xlab = "Temperature")
hist(pc_df$Humidity, main="Histogram of Humidity", col = "blue", xlab = "Humidity")
hist(pc_df$`Wind Speed`, main="Histogram of Wind Speed", col = "blue", xlab = "Wind Speed")
hist(pc_df$`general diffuse flows`, main="Histogram of General Diffuse", col = "blue", xlab = "General Diffuse Flow")
hist(pc_df$`diffuse flows`, main="Histogram of Diffuse", col = "blue", xlab ="Diffuse Flow")
hist(pc_df$`Zone 1 Power Consumption`, main="Histogram of Zone 1", col = "blue", xlab ="Zone 1 Power Consumption")
hist(pc_df$`Zone 2  Power Consumption`, main="Histogram of Zone 2", col = "blue", xlab ="Zone 2 Power Consumption")
hist(pc_df$`Zone 3  Power Consumption`, main="Histogram of Zone 3", col = "blue", xlab ="Zone 3 Power Consumption")
```
Temperature appers to have a normal distribution. Humidity is left skewed- as mentioned in summary statistics, humidity is pretty high. Wind Speed is interesting as well, looks like either little to low wind speeds or ~4.5 km/h wind speed . General and Diffuse flows are right skewed. Flow tends to be on a lower level. Zone 1 and Zone 2 have a prety normal distribution. Zone 2 looks normal as well but looks a bit right skewed. 

```{r}
plot(pc_df$`Zone 1 Power Consumption`,
     type="l",
     col =2,
     ylim = c(5000,55000),
     xlab = "Date",
     ylab = "Power Consumption (kW)")
lines(pc_df$`Zone 2  Power Consumption`,
       type="l",
     col =3)
lines(pc_df$`Zone 3  Power Consumption`,
       type="l",
     col =4)
legend("topright", c("Zone 1", "Zone 2", "Zone 3"), lty = 1, col = 2:4)
```
Looking at all three zone's power consumption, it looks like Zone 1 has more power consumption than both Zone 2 and Zone 3. 


Grouping Time Series To look at Daily:
```{r}
#Creating Column with just the date
pc_df$Date <- as.Date(pc_df$DateTime, format="%m/%d/%Y")
head(pc_df)

#Combining Date: 
pc_zone1_daily <- pc_df %>% group_by(Date) %>% summarise(Total= sum(`Zone 1 Power Consumption`))
pc_zone2_daily <- pc_df %>% group_by(Date) %>% summarise(Total= sum(`Zone 2  Power Consumption`))
pc_zone3_daily <- pc_df %>% group_by(Date) %>% summarise(Total= sum(`Zone 3  Power Consumption`))

head(pc_zone1_daily)
head(pc_zone2_daily)
head(pc_zone3_daily)

#Setting as time series:

pc_ts1_daily  <- ts(pc_zone1_daily$Total)
pc_ts2_daily  <- ts(pc_zone2_daily$Total)
pc_ts3_daily  <- ts(pc_zone3_daily$Total)
```

Plotting Time Series Daily:
```{r}
#Daily Time Series Zone 1
autoplot(pc_ts1_daily) +
  labs(title = "Zone 1 Daily Power Consumption",
       x = "Day",
       y = "Power Consumption (kW)") +
  theme_classic()

#Daily Time Series Zone 2
autoplot(pc_ts2_daily) +
  labs(title = "Zone 2 Daily Power Consumption",
       x = "Day",
       y = "Power Consumption (kW)") +
  theme_classic()

#Daily Time Series Zone 3
autoplot(pc_ts3_daily) +
  labs(title = "Zone 3 Daily Power Consumption",
       x = "Day",
       y = "Power Consumption (kW)") +
  theme_classic()

```
All Three zones have similar trending path- Power consumption increases up until it reaches a peak followed by a decline. Looks like there's some seasonlity.There are some sharp declines on random days- perhaps indicators of power outages or something

```{r}
#Train Test Split

X <- pc_df[ , c("Temperature", "Humidity", "Wind Speed", "general diffuse flows", "diffuse flows", "Date")] 
head(X)

y1 <- pc_df$`Zone 1 Power Consumption`
y2 <- pc_df$`Zone 2  Power Consumption`
y3 <- pc_df$`Zone 3  Power Consumption`

#Zone 1 Splitting:
nTotal1 <- length(y1)
nValid1 <- 75
nTrain1 <- nTotal1 - nValid1
xTrain1 <- X[1:nTrain1, ]
yTrain1 <- y1[1:nTrain1]
xValid1 <- X[(nTrain1 + 1):nTotal1, ]
yValid1 <- y1[(nTrain1 + 1):nTotal1]

yTrain1.ts <- ts(yTrain1, start = c(2017, 1), end= c(2017, 289), frequency = 365)
yValid1.ts <- ts(yTrain1, start = c(2017, 290), end= c(2017, 364),frequency = 365)

#Zone 2 Splitting:
nTotal2 <- length(y2)
nValid2 <- 75
nTrain2 <- nTotal2 - nValid2
xTrain2 <- X[1:nTrain2, ]
yTrain2 <- y2[1:nTrain2]
xValid2 <- X[(nTrain1 + 1):nTotal2, ]
yValid2 <- y2[(nTrain1 + 1):nTotal2]


yTrain2.ts <- ts(yTrain2, start = c(2017, 1), end= c(2017, 289), frequency = 365)
yValid2.ts <- ts(yTrain2, start = c(2017, 290), end= c(2017, 364),frequency = 365)

#Zone 3 Splitting:
nTotal3 <- length(y3)
nValid3 <- 75
nTrain3 <- nTotal3 - nValid3
xTrain3 <- X[1:nTrain3, ]
yTrain3 <- y3[1:nTrain3]
xValid3 <- X[(nTrain1 + 1):nTotal3, ]
yValid2 <- y3[(nTrain1 + 1):nTotal3]


yTrain3.ts <- ts(yTrain3, start = c(2017, 1), end= c(2017, 289), frequency = 365)
yValid3.ts <- ts(yTrain3, start = c(2017, 290), end= c(2017, 364),frequency = 365)
```
Modeling Zone 1:

Holt Winters:
```{r}
#Holt Winters Exponential Smoothing

#Zone 1
pc1_holt.model <- ets(yTrain1.ts, model = "ZAA", alpha = .2, gamma = .05)

pc1_holt.forecast <- forecast(pc1_holt.model, h = 75, level = c(.95))

accuracy(pc1_holt.forecast, yValid1.ts)

#plot the forecast
autoplot(yTrain1.ts, color = "black") + 
  autolayer(pc1_holt.forecast, color = "red") +
  autolayer(yValid1.ts, color = "blue") +
  labs(title = "Zone 1 Power Consumption Forecast- Holt Winters",
       x = "Date",
       y = "Power Consumption (kW)")+
  theme_classic()
```

Linear Modeling:
```{r}
# Linear:
pc1_lm <- tslm(yTrain1.ts ~ trend)

pc1_lm.forecast <- forecast(pc1_lm, h = nValid1, level=0)

accuracy(pc1_lm.forecast, yValid1.ts)

#plot the forecast
autoplot(yTrain1.ts, color = "black") + 
  autolayer(pc1_lm.forecast, color = "red") +
  autolayer(yValid1.ts, color = "blue") +
  labs(title = "Zone 1 Power Consumption Forecast - Linear Model",
       x = "Date",
       y = "Power Consumption (kW)")+
  theme_classic()
```
```{r}
#Modeling with exponential trend:
pc1_expo.trend <- tslm(yTrain1.ts ~ trend, lambda = 1)
pc1_expo.forecast <- forecast(pc1_expo.trend, h =nValid1, level=0)

accuracy(pc1_expo.forecast, yValid1.ts)

#plot the forecast
autoplot(yTrain1.ts, color = "black") + 
  autolayer(pc1_expo.forecast, color = "red") +
  autolayer(yValid1.ts, color = "blue") +
  labs(title = "Zone 1 Power Consumption Forecast - Expo Model",
       x = "Date",
       y = "Power Consumption (kW)")+
  theme_classic()

```

