---
title: "Final Team Project - Team 4"
author: "Uyen Pham, Jacqueline Urenda, and Renetta Nelson"
date: "December 5, 2022"
output: pdf_document
---



# Loading Libraries

```{r, message=FALSE}

library(tidyverse)
library(fpp2)
library(readr)
library(forecast)
library(ggplot2)
library(gridExtra)
library(reshape2)
library(dplyr)
library(lubridate)
library(RColorBrewer)
library(corrplot)
set.seed(506)


```

# Loading the Dataset 

```{r}

#Load the data 
power <- read_csv("Tetuan City power consumption.csv", 
    col_types = cols(DateTime = col_datetime(format = "%m/%d/%Y %H:%M")))

```

# Data Transformation

The columns in the data set are renamed. This was done for easy reference to variables.

```{r}

colnames(power) <- c('DateTime', 'Temperature', 'Humidity','Wind_Speed', 
                     'Gen_Diffuse_Flows', 'Diffuse_Flows', 'Zone1', 'Zone2', 
                     'Zone3')
head(power)

print("Missing Values: ")
sum(is.na(power)) #no missing values

```


# Statistical Data Analysis

The data set consists of 364 days total, taking data from January 1, 2017 to December 30, 2017. The time window is every ten minutes. The temperature is measured in Celsius with a mean and median around 18 degrees. The humidity column displays the percentage of the humidity. The average humidity is 68.26%. The wind speed is measured in km/h, and the power consumption is measured in KiloWatts.


```{r}

summary(power)

```


# Exploratory Data Analysis


## Histograms

Temperature appears to have a normal distribution. Humidity is left skewed- as mentioned in summary statistics, humidity is pretty high. Wind Speed is interesting as well, looks like either little to low wind speeds or ~4.5 km/h wind speed . General and Diffuse flows are right skewed. Flow tends to be on a lower level. Zone 1 and Zone 2 have a normal distribution. Zone 2 looks normal as well but looks a bit right skewed. 

```{r}
par(mfrow=c(3,3))
hist(power$Temperature, main="Histogram of Temperature", col = "blue",
     xlab = "Temperature")
hist(power$Humidity, main="Histogram of Humidity", col = "blue",
     xlab = "Humidity")
hist(power$Wind_Speed, main="Histogram of Wind Speed", col = "blue",
     xlab = "Wind Speed")
hist(power$Gen_Diffuse_Flows, main="Histogram of General Diffuse",
     col = "blue", xlab = "General Diffuse Flow")
hist(power$Diffuse_Flows, main="Histogram of Diffuse", col = "blue",
     xlab ="Diffuse Flow")
hist(power$Zone1, main="Histogram of Zone 1", col = "blue",
     xlab ="Zone 1 Power Consumption")
hist(power$Zone2, main="Histogram of Zone 2", col = "blue", 
     xlab ="Zone 2 Power Consumption")
hist(power$Zone3, main="Histogram of Zone 3", col = "blue", 
     xlab ="Zone 3 Power Consumption")
```

## Comparisons of the Power Consumption Zones

Looking at all three zone's power consumption, it looks like Zone 1 has more power consumption than both Zone 2 and Zone 3. 

```{r}
plot(power$Zone1,
     type="l",
     col =2,
     ylim = c(5000,55000),
     xlab = "Date",
     ylab = "Power Consumption (kW)")
lines(power$Zone2,
       type="l",
     col =3)
lines(power$Zone3,
       type="l",
     col =4)
legend("topright", c("Zone 1", "Zone 2", "Zone 3"), lty = 1, col = 2:4)
```


## Correlations

All three zones have strong correlations between each other. Temperature and wind speed have a correlation of 0.48. Humidity and temperature have a moderate but noticeable correlation as well. General diffuse flows and Diffuse flows have a correlation of 0.56. Temperature seems to show highest correlation with power consumption.


```{r}

corr <- round(x = cor(power[, 2:9]), digits = 2)

corrplot(corr, method = "color", outline = T, addgrid.col = "darkgray", 
         order="hclust", addrect = 4, rect.col = "black", rect.lwd = 5,
         cl.pos = "b", tl.col = "indianred4", tl.cex = .75, cl.cex = .75,
         addCoef.col = "white", number.digits = 2, number.cex = 0.5, col =
           colorRampPalette(c("darkred","white","midnightblue"))(100))
```


# Monthly Power Consumption

It is more interesting to explore the daily data, since we would be more interesting to forecast the power each Zone needed daily, so the power providers could increase or decrease power supply to meet the demand.


```{r Monthly}

power_monthly <- power %>% group_by(Month = format(DateTime, "%Y - %m")) %>% 
  summarise(Total = sum(Zone1))

ggplot(power_monthly, aes(x = Month, y = Total, group = 1)) + geom_line(size = 1) +
  theme_classic()

```


# Daily Power Consumption

All three zones for power consumption have similar trending paths. The power consumption increases up until it reaches a peak followed by a decline. There appears to be some seasonality. There are some sharp declines on random days, perhaps indicating a presence of power outages. Zone 1 and Zone 2 seem to have big daily fluctuation while Zone 3 seem less fluctuated. Also, They all seem to peak at around May to September which could be due to more power consumption during Summer time.

```{r Daily}
power$Date <- as.Date(power$DateTime, format = "%m/%d/%Y")
power.daily.z1 <- power %>% group_by(Date) %>% summarise(Total= sum(Zone1))
power.daily.z2 <- power %>% group_by(Date) %>% summarise(Total= sum(Zone2))
power.daily.z3 <- power %>% group_by(Date) %>% summarise(Total= sum(Zone3))

p1<- ggplot(power.daily.z1, aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Daily Power Consumption in Zone 1",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))

p2 <- ggplot(power.daily.z2, aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Daily Power Consumption in Zone 2",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))

p3 <- ggplot(power.daily.z3, aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Daily Power Consumption in Zone 3",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))
grid.arrange(p1, p2, p3, ncol = 2)
```

# Mean Power Consumption

```{r}

#Mean model

power.daily.z1.ts1 <-ts(power.daily.z1$Total) 

mean_power = mean(power.daily.z1$Total)

autoplot(power.daily.z1.ts1) + labs( title = "Mean Model of Power Consumption",
                                     x = "Daily Time", y = "Power Consumption")+
  geom_hline(yintercept = mean_power, color = "green", size = 1) + theme_classic()



```


# Random Walk

Autocorrelation show strong correlation at multiple lag showing weekly seasonality (~0.02, 0.04, 0.06 of 365 days).

```{r}
#Convert daily power in 3 zones into time series object
power.daily.z1.ts <- ts(power.daily.z1$Total)
power.daily.z2.ts <- ts(power.daily.z2$Total)
power.daily.z3.ts <- ts(power.daily.z3$Total)

#Compute and plot the autocorrelation and produce an ACF plot
par(mar=c(5,5,5,2)) #To adjust the plot size and not get the title cut off
acf_value1 <- Acf(power.daily.z1.ts, lag.max=1, 
                  main='Zone 1 Autocorrelation Plot with Lag-1')
acf_value2 <- Acf(power.daily.z2.ts, lag.max=1, 
                  main='Zone 2 Autocorrelation Plot with Lag-1')
acf_value3 <- Acf(power.daily.z3.ts, lag.max=1, 
                  main='Zone 3 Autocorrelation Plot with Lag-1')
```

```{r}
print(acf_value1)
print(acf_value2)
print(acf_value3)
```


```{r}
#compute and plot at the autocorrelation of the daily data
par(mar=c(5,5,5,2))
acf_value <- acf(power.daily.z1.ts)
#acf_value <- acf(zone1.min.ts)
```


# First Three Weeks of the Different Zones

The following shows the first few weeks of power consumption for all the zones. Zone 1 and 2 starts low on Sunday, goes high during the weekdays, and then low again on Saturday. Zone 3 seems to be have the opposite behavior and the difference between days are less pronounced. 

```{r zoom in zone 1}
#Zoom in the first 3 weeks of Zone 1
ggplot(power.daily.z1[0:21, ], aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Zone 1 Power Consumption",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))
  #coord_cartesian(xlim = c(1, 14))
```



```{r zoom in zone 2}
#Zoom in the first 3 weeks of Zone 1
ggplot(power.daily.z2[0:21, ], aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Zone 2 Power Consumption",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))
  #coord_cartesian(xlim = c(1, 14))
```


```{r zoom in zone 3}
#Zoom in the first 3 weeks of Zone 3
ggplot(power.daily.z3[0:21, ], aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Zone 3 Power Consumption",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))
  #coord_cartesian(xlim = c(1, 14))
```

#Partitioning the Data


```{r Create new variables and dataframes}

#Convert power data frame to daily data including all the variables
daily <- power[ ,2:10] %>% group_by(Date) %>% 
  summarise(across(everything(), sum), .groups = 'drop')

#Create day of the week column
daily$DOW <- wday(daily$Date, label = TRUE)

#Create month column 
daily$Month <- month(daily$Date, label = TRUE)

#Create new frames with values in Month, DOW converted to dummies
DOW.dummies <- model.matrix(~ 0 + DOW, data = daily)

#Rename each dummies column without "Month", "Dow" in front 
colnames(DOW.dummies) <- gsub("DOW", "", colnames(DOW.dummies))

#Combine the data frames
X <- as.data.frame(cbind(daily[,2:6], DOW.dummies[, -3]))

#Create y series with power values from daily data frame for zone 1
y1 <- daily$Zone1

#Split data 
nTotal <- length(y1) 
nValid <- 56
nTrain <- nTotal - nValid
xTrain <- X[1:nTrain, ]
xValid <- X[(nTrain + 1):nTotal, ]
yTrain1 <- y1[1:nTrain] # Zone 1
yValid1 <- y1[(nTrain + 1):nTotal]

#Convert y train and validation into time series object
yTrain1.ts <- ts(yTrain1, start = c(1, 1), end= c(44, 7), frequency = 7)
yValid1.ts <- ts(yValid1, start = c(45, 1), end= c(52, 7), frequency = 7)

autoplot(yTrain1.ts ) + autolayer(yValid1.ts, color = "blue") + theme_classic() + 
  labs(title = "Training and Validation Sets", y = "Power Consumption (kW)")

```

# Decomposition of the Time Series

```{r}

stl.run <- stl(yTrain1.ts, s.window = "periodic")
plot(stl.run)
```

# Modeling

## Naive Forecast

```{r}

naive <- naive(yTrain1.ts, h=nValid)
accuracy(naive, yValid1.ts)


autoplot(yTrain1.ts, color = "black") + 
  autolayer(naive, color = "red") +
  autolayer(yValid1.ts, color = "blue") +
  labs(title = "Zone 1 Power Consumption Forecast- Naive",
       x = "Date",
       y = "Power Consumption (kW)")+
  theme_classic()


```


## Seasonal Naive Forecast

```{r}

sn_model <- snaive(yTrain1.ts, h=nValid, level = 95)

autoplot(yTrain1.ts) + autolayer(sn_model, color = "#006d2c", 
                                 series = "Seasonal Naive") + 
  autolayer(yValid1.ts, color = "blue", series = "Actual") + theme_classic() + 
  labs(title = "Zone 1 Power Consumption Forecast - Seasonal Naive",
       y = "Power Consumption (kW)")


```


## Holt Winters

```{r}
#Holt Winters Exponential Smoothing

#Zone 1
pc1_holt.model <- ets(yTrain1.ts, model = "ZAA", alpha = .2, gamma = .05)

pc1_holt.forecast <- forecast(pc1_holt.model, h = nValid, level = c(.95))

accuracy(pc1_holt.forecast, yValid1.ts)

#plot the forecast
autoplot(yTrain1.ts, color = "black") + 
  autolayer(pc1_holt.forecast, color = "red") +
  autolayer(yValid1.ts, color = "blue") +
  labs(title = "Zone 1 Power Consumption Forecast- Holt Winters",
       x = "Date",
       y = "Power Consumption (kW)")+
  theme_classic()

```

## TLSM 

```{r TSLM model}
#Create the formula for the regression model
(formula <- as.formula(paste("yTrain1.ts", paste(c("trend", colnames(xTrain)),
                                                 collapse = "+"), sep = "~")))
#Train the regression model
zone1.tslm <- tslm(formula, data = xTrain) #lamda=1 for linear trend
#Predict valid set
zone1.tslm.pred <- forecast(zone1.tslm, newdata = xValid)
#Plot the prediction
plot(zone1.tslm.pred, xlab = "Days", ylab = "Zone 1 Power Consumption")

summary(zone1.tslm)

accuracy(zone1.tslm.pred, yValid1.ts)
```


```{r}
#Create the formula for the regression model

#Train the regression model
zone1.tslm.test <- tslm(yTrain1.ts ~ trend) #lamda=1 for linear trend
#Predict valid set
zone1.tslm.test.pred <- forecast(zone1.tslm.test, newdata = xValid)
#Plot the prediction
plot(zone1.tslm.test.pred, xlab = "Days", ylab = "Zone 1 Power Consumption")

accuracy(zone1.tslm.test.pred, yValid1.ts)

```


```{r}
tslm(yTrain1.ts ~ trend + I(trend^2) + I(sin(2*pi*trend/12))
+ I(cos(2*pi*trend/12)))
```

## Linear Model

```{r}

pc1_lm <- tslm(yTrain1.ts ~ trend)

pc1_lm.forecast <- forecast(pc1_lm, h = nValid, level=0)

accuracy(pc1_lm.forecast, yValid1.ts)

#plot the forecast
autoplot(yTrain1.ts, color = "black") + 
  autolayer(pc1_lm.forecast, color = "red") +
  autolayer(yValid1.ts, color = "blue") +
  labs(title = "Zone 1 Power Consumption Forecast - Linear Model",
       x = "Date",
       y = "Power Consumption (kW)")+
  theme_classic()
```

## Exponential Trend

```{r}

pc1_expo.trend <- tslm(yTrain1.ts ~ trend, lambda = 1)
pc1_expo.forecast <- forecast(pc1_expo.trend, h =nValid, level=0)

accuracy(pc1_expo.forecast, yValid1.ts)

#plot the forecast
autoplot(yTrain1.ts, color = "black") + 
  autolayer(pc1_expo.forecast, color = "red") +
  autolayer(yValid1.ts, color = "blue") +
  labs(title = "Zone 1 Power Consumption Forecast - Expo Model",
       x = "Date",
       y = "Power Consumption (kW)")+
  theme_classic()

```

## Neural Network

```{r}

pc1.nnetar <- nnetar(yTrain1.ts)
summary(pc1.nnetar$model[[1]])
pc.1.nnetar.pred <- forecast(pc1.nnetar, h=nValid)

accuracy(pc.1.nnetar.pred, yValid1.ts)


#plot the forecast
autoplot(yTrain1.ts, color = "black") + 
  autolayer(pc.1.nnetar.pred, color = "red") +
  autolayer(yValid1.ts, color = "blue") +
  labs(title = "Zone 1 Power Consumption Forecast - Neural Network",
       x = "Date",
       y = "Power Consumption (kW)")+
  theme_classic()
```

## Arima Model

```{r}
auto_model <- auto.arima(yTrain1.ts)
auto_model


par(mar=c(5,5,5,2))
pacf(yTrain1.ts, lag.max =52)
acf(yTrain1.ts, lag.max = 52)
```




```{r}
#Take the difference for the training set and observe the autoregression
yTrain1.lag1.ts <- diff(diff(diff(diff(yTrain1.ts, lag = 7))))
plot(yTrain1.lag1.ts)

par(mar=c(5,5,5,2))
pacf(yTrain1.lag1.ts, lag.max =52)
acf(yTrain1.lag1.ts, lag.max = 52)

```

```{r my Arima}
predictor_train <- as.matrix(xTrain)
#Run Arima model
zone1.arima.model <- Arima(yTrain1, order = c(1,3,1), xreg =predictor_train )
summary(zone1.arima.model)
#Forecast on the test set
predictors_test <- as.matrix(xValid)
zone1.arima_predict <- forecast(zone1.arima.model, h = nValid, 
                                xreg= predictors_test)
zone1.arima_predict.ts <- ts(zone1.arima_predict$mean, start = c(45, 1),
                             end= c(52,7), frequency = 7)

accuracy(zone1.arima_predict.ts, yValid1.ts)

```


```{r}
#plot the forecast
autoplot(yTrain1.ts) + 
  autolayer(zone1.arima_predict.ts, color = "red") +
  autolayer(yValid1.ts, color = "gray") +
  labs(title = "Zone 1 Power Consumption Forecast - ARIMA ",
       x = "Date",
       y = "Power Consumption (kW)")+
  theme(plot.title=element_text(hjust=0.5)) 
  #+coord_cartesian(xlim = c(, ))
```












