---
title: "Final Team Project - Team 4"
author: "Uyen Pham, Jacqueline Urenda, and Renetta Nelson"
date: "December 5, 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r, message=FALSE}
library(tidyverse)
library(fpp2)
library(readr)
library(forecast)
library(ggplot2)
library(gridExtra)
library(reshape2)
library(dplyr)
library(lubridate)
library(RColorBrewer)
library(corrplot)
library(Hmisc)
library(ggpubr)
set.seed(506)
```

## Loading the Dataset 

```{r}
#Load the data 
power <- read_csv("Tetuan City power consumption.csv", 
    col_types = cols(DateTime = col_datetime(format = "%m/%d/%Y %H:%M")))
```


The columns in the data set are renamed. This was done for easy reference to variables.

```{r}
colnames(power) <- c('DateTime', 'Temperature', 'Humidity','Wind_Speed', 
                     'Gen_Diffuse_Flows', 'Diffuse_Flows', 'Zone1', 'Zone2', 
                     'Zone3')
head(power)
```


## Exploratory Data Analysis

```{r}
print("Missing Values: ")
sum(is.na(power)) #no missing values
```


### Statistical Data Analysis

The data set consists of 364 days total, taking data from January 1, 2017 to December 30, 2017. The time window is every ten minutes. The temperature is measured in Celsius with a mean and median around 18 degrees. The humidity column displays the percentage of the humidity. The average humidity is 68.26%. The wind speed is measured in km/h, and the power consumption is measured in KiloWatts.


```{r}
summary(power)
```


### Histograms

Temperature appears to have a normal distribution. Humidity is left skewed- as mentioned in summary statistics, humidity is pretty high. Wind Speed is interesting as well, looks like either little to low wind speeds or ~4.5 km/h wind speed . General and Diffuse flows are right skewed. Flow tends to be on a lower level. Zone 1 and Zone 2 have a normal distribution. Zone 2 looks normal as well but looks a bit right skewed. 


```{r}
#Histograms of all variables
hist.data.frame(power[,2:9])
```

### Correlations

All three zones have strong correlations between each other. Temperature and wind speed have a correlation of 0.48. Humidity and temperature have a moderate but noticeable correlation as well. General diffuse flows and Diffuse flows have a correlation of 0.56. Temperature seems to show highest correlation with power consumption.


```{r}
corr <- round(x = cor(power[, 2:9]), digits = 2)
corrplot(corr, method = "color", outline = T, addgrid.col = "darkgray", 
         order="hclust", addrect = 4, rect.col = "black", rect.lwd = 5,
         cl.pos = "b", tl.col = "indianred4", tl.cex = .75, cl.cex = .75,
         addCoef.col = "white", number.digits = 2, number.cex = 0.5, col =
           colorRampPalette(c("darkred","white","midnightblue"))(100))
```

```{r}
#Observe the the pair plots between variables
#plot(power[, 2:9],  pch = 16, cex= 0.3)
```

### Comparisons of the Power Consumption Zones

Looking at all three zone's power consumption, it looks like Zone 1 has more power consumption than both Zone 2 and Zone 3. 

```{r}
plot(power$Zone1,
     type="l",
     col =2,
     ylim = c(5000,55000),
     xlab = "Number of 10-minute interval",
     ylab = "Power Consumption (kW)")
lines(power$Zone2,
       type="l",
     col =3)
lines(power$Zone3,
       type="l",
     col =4)
legend("topright", c("Zone 1", "Zone 2", "Zone 3"), lty = 1, col = 2:4)
```

# Explore Power Consumption in different time intervals

```{r}
#Hourly Power consumption
power_hourly <- power %>%
 group_by(Hour= format(DateTime, "%Y-%m-%d %H")) %>%
 summarise(Total= sum(Zone1))

  #Convert Zone1_Consumption to time series
hour.zone1.ts <- ts(power_hourly$Total, start= c(2017,1), end=c(2017, 8376), frequency=8376)
  #Plot the ts
p1 <- autoplot(hour.zone1.ts, color="blue", main = "Hourly Power Consumption", cex =0.3)

#Daily  Power consumption
power$Date <- as.Date(power$DateTime, format = "%m/%d/%Y")
power.daily.z1 <- power %>% group_by(Date) %>% summarise(Total= sum(Zone1))

p2 <- ggplot(power.daily.z1, aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Daily Power Consumption in Zone 1",
       x = "Day",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.3, size=10))

#Weekly Power consumption
power_weekly <- power %>%
 group_by(week = lubridate::week(Date)) %>% summarise(Total= sum(Zone1))

p3 <- ggplot(power_weekly, aes(x = week, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Weekly Power Consumption in Zone 1",
       x = "Week",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.3, size=10))

#Monthly Power consumption
power_monthly <- power %>%
  group_by(Month = format(Date, "%Y-%m")) %>%
  summarise(Total = sum(Zone1))

p4 <- ggplot(power_monthly, aes(x = Month, y = Total, group = 1)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Monthly Power Consumption in Zone 1",
       x = "Month",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.3, size=10))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
grid.arrange(p1, p2, p3, p4, ncol = 2)
```

It is more interesting to explore the daily data, since we would be more interesting to forecast the power each Zone needed daily, so the power providers could increase or decrease power supply to meet the demand.

### Daily Power Consumption for all three zones

All three zones for power consumption have similar trending paths. The power consumption increases up until it reaches a peak followed by a decline. There appears to be some seasonality. There are some sharp declines on random days, perhaps indicating a presence of power outages. Zone 1 and Zone 2 seem to have big daily fluctuation while Zone 3 seem less fluctuated. Also, They all seem to peak at around May to September which could be due to more power consumption during Summer time.

```{r Daily}
power$Date <- as.Date(power$DateTime, format = "%m/%d/%Y")
power.daily.z1 <- power %>% group_by(Date) %>% summarise(Total= sum(Zone1))
power.daily.z2 <- power %>% group_by(Date) %>% summarise(Total= sum(Zone2))
power.daily.z3 <- power %>% group_by(Date) %>% summarise(Total= sum(Zone3))
p1<- ggplot(power.daily.z1, aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Daily Power Consumption in Zone 1",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))
p2 <- ggplot(power.daily.z2, aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Daily Power Consumption in Zone 2",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))
p3 <- ggplot(power.daily.z3, aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Daily Power Consumption in Zone 3",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))
grid.arrange(p1, p2, p3, ncol = 1)
```

### Check for random walk 

```{r}
#Convert daily power in 3 zones into time series object
power.daily.z1.ts <- ts(power.daily.z1$Total)
power.daily.z2.ts <- ts(power.daily.z2$Total)
power.daily.z3.ts <- ts(power.daily.z3$Total)
```


Autocorrelation show high lag 1 correlation in all three zones with zone 1 very close to random walk while zone 2 and 3 also show correlation at multiple lags which could be weekly seasonality. 

```{r acf}
#Compute and plot the autocorrelation and produce an ACF plot
par(mfrow=c(2,2))
par(mar=c(5,5,5,2))
#To adjust the plot size and not get the title cut off
acf_value1 <- acf(power.daily.z1.ts, lag.max=364, 
                  main='Zone 1 Autocorrelation Plot with Lag-1')
acf_value2 <- acf(power.daily.z2.ts, lag.max=364, 
                  main='Zone 2 Autocorrelation Plot with Lag-1')
acf_value3 <- acf(power.daily.z3.ts, lag.max=364, 
                  main='Zone 3 Autocorrelation Plot with Lag-1')
```

```{r}
#Get value at lag-1
print(acf_value1[1])
print(acf_value2[1])
print(acf_value3[1])
```


### Examine the first Three Weeks of the Different Zones

The following shows the first few weeks of power consumption for all the zones. Zone 1 and 2 starts low on Sunday, goes high during the weekdays, and then low again on Saturday. Zone 3 seems to be have the opposite behavior and the difference between days are less pronounced. 

```{r zoom in the first 3 weeks}
#Zoom in the first 3 weeks of Zone 1
p1 <- ggplot(power.daily.z1[0:21, ], aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Zone 1 Power Consumption",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))

#Zoom in the first 3 weeks of Zone 1
p2 <- ggplot(power.daily.z2[0:21, ], aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Zone 2 Power Consumption",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))

#Zoom in the first 3 weeks of Zone 3
p3 <- ggplot(power.daily.z3[0:21, ], aes(x = Date, y = Total)) +
  geom_line(size = 0.5, color = "blue") +
  labs(title = "Zone 3 Power Consumption",
       x = "Daily",
       y = "Power Consumption") +
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))
grid.arrange(p1, p2, p3, ncol = 1)
```

## Partitioning the Data


```{r}
#Convert power data frame to daily data including all the variables
daily.sum <- power[ ,2:10] %>%
   group_by(Date) %>%
   summarise(across(everything(), sum), .groups = 'drop')
#Compute daily mean of the variables (Temperature, Humidity, Wind_Speed, Gen_Diffuse_Flows, Diffuse_Flows) 
daily.var.mean <- (daily.sum[ ,2:6])/144
#Create day of the week column
daily.sum$DOW <- wday(daily.sum$Date, label = TRUE)
#create new frames with values DOW converted to dummies
DOW.dummies <- model.matrix(~ 0 + DOW, data = daily.sum)
#Rename each dummies column without  "Dow" in front 
colnames(DOW.dummies) <- gsub("DOW", "", colnames(DOW.dummies))
#Combine the data frames, exclude Tues to avoid dummy trap
X <- as.data.frame(cbind(daily.var.mean, DOW.dummies[, -3]))

#Create y series with power values from daily data frame for zone 1
y1 <- daily.sum$Zone1
y2 <- daily.sum$Zone2
y3 <- daily.sum$Zone3
#Split data
nTotal <- length(y1) #could use length zone 1 for all 3 zones
nValid <- 56
nTrain <- nTotal - nValid
xTrain <- X[1:nTrain, ]
xValid <- X[(nTrain + 1):nTotal, ]
# 
##Response variable for train and valid sets
    # Zone 1
yTrain1 <- y1[1:nTrain]
yValid1 <- y1[(nTrain + 1):nTotal]

yTrain2 <- y2[1:nTrain]
yValid2 <- y2[(nTrain + 1):nTotal]

yTrain3 <- y3[1:nTrain]
yValid3 <- y3[(nTrain + 1):nTotal]

#Convert y train and validation into time series object
yTrain1.ts <- ts(yTrain1, start = c(1, 1), end= c(44, 7), frequency = 7)
yValid1.ts <- ts(yValid1, start = c(45, 1), end= c(52, 7), frequency = 7)

yTrain2.ts <- ts(yTrain2, start = c(1, 1), end= c(44, 7), frequency = 7)
yValid2.ts <- ts(yValid2, start = c(45, 1), end= c(52, 7), frequency = 7)

yTrain3.ts <- ts(yTrain3, start = c(1, 1), end= c(44, 7), frequency = 7)
yValid3.ts <- ts(yValid3, start = c(45, 1), end= c(52, 7), frequency = 7)
```

# Decomposition of the Time Series 

```{r}
#Decompose time series
stl1.run <- stl(yTrain1.ts, s.window = "periodic")
stl2.run <- stl(yTrain2.ts, s.window = "periodic")
stl3.run <- stl(yTrain3.ts, s.window = "periodic")

plot(stl1.run)
plot(stl2.run)
plot(stl3.run)
```

```{r}
# autoplot(yTrain1.ts ) + autolayer(yValid1.ts, color = "blue") + theme_classic() + 
#   labs(title = "Training and Validation Sets", y = "Power Consumption (kW)")
# autoplot(yTrain2.ts ) + autolayer(yValid2.ts, color = "blue") + theme_classic() + 
#   labs(title = "Training and Validation Sets", y = "Power Consumption (kW)")
# autoplot(yTrain3.ts ) + autolayer(yValid3.ts, color = "blue") + theme_classic() + 
#   labs(title = "Training and Validation Sets", y = "Power Consumption (kW)")
```


# Modeling

## Mean Model

```{r}
#Average model function 
mean_predict <- function(yTrain.ts, yValid.ts, titl) {
  mean_power = meanf(yTrain.ts, h=nValid )
  mean_power.ts <- ts( mean_power$mean, start = c(45, 1), end= c(52, 7), 
                       frequency = 7)
  p <- autoplot(yTrain.ts) + 
    autolayer(mean_power$mean, color = 'red')+
    autolayer(yValid.ts, color = "blue")+
    labs(title =titl,x = "week", y = "Power Consumption")+
    theme (title =element_text(size=7))+
    geom_hline(yintercept = mean_power$mean, color = "green", size = 0.5)
  acc <- accuracy(mean_power$mean, yValid.ts)
  lst <- list(acc,p)
  return(lst)}
#call model output
zone1.mean<- mean_predict(yTrain1.ts, yValid1.ts, "Zone 1")
zone2.mean<- mean_predict(yTrain2.ts, yValid2.ts, "Zone 2")
zone3.mean<- mean_predict(yTrain3.ts, yValid3.ts, "Zone 3")
```

```{r}
zone1.mean[[1]]
zone2.mean[[1]]
zone3.mean[[1]]
```


```{r}
ggarrange(zone1.mean[[2]], zone2.mean[[2]], zone3.mean[[2]], ncol=1)
```

## Naive Forecast

```{r}
# Naive Forecast function for all 3 zones
naive <- function(yTrain.ts, yValid.ts, titl) {
  naive_model = forecast::naive(yTrain.ts, h=nValid)
  #naive_model.pred<- forecast(naive_model, h = nValid, level = c(.95))
  naive_model.ts <- ts(naive_model$mean, start = c(45, 1), end= c(52, 7), 
                            frequency = 7)
  p <- autoplot(yTrain.ts) + 
    autolayer(naive_model$mean, color = 'red')+
    autolayer(yValid.ts, color = "blue")+
    labs(title =titl,x = "week", y = "Power Consumption")+
    theme (title =element_text(size=7))
    #geom_hline(yintercept = (naive_model$mean, color = "green", size = 0.5))
  acc <- accuracy(naive_model$mean, yValid.ts)
  lst <- list(acc,p)
  return(lst)}
#call model output
zone1.naive<- naive(yTrain1.ts, yValid1.ts, "Zone 1")
zone2.naive<- naive(yTrain2.ts, yValid2.ts, "Zone 2")
zone3.naive<- naive(yTrain3.ts, yValid3.ts, "Zone 3")
```

```{r}
zone1.naive[[1]]
zone2.naive[[1]]
zone3.naive[[1]]
```


```{r}
ggarrange(zone1.naive[[2]], zone2.naive[[2]], zone3.naive[[2]], ncol=1)
```

## Seasonal Naive Forecast

```{r}
# Build seasonal Naive Forecast function for all 3 zones
snaive <- function(yTrain.ts, yValid.ts, titl) {
  snaive_model = forecast::snaive(yTrain.ts, h=nValid)
  snaive_model.ts <- ts(snaive_model$mean, start = c(45, 1), end= c(52, 7), 
                            frequency = 7)
  p <- autoplot(yTrain.ts) + 
    autolayer(snaive_model$mean, color = 'red')+
    autolayer(yValid.ts, color = "blue")+
    labs(title =titl,x = "week", y = "Power Consumption")+
    theme (title =element_text(size=7))
    #geom_hline(yintercept = (naive_model$mean, color = "green", size = 0.5))
  acc <- accuracy(snaive_model$mean, yValid.ts)
  lst <- list(acc,p)
  return(lst)}
#call model output
zone1.snaive<- snaive(yTrain1.ts, yValid1.ts, "Zone 1")
zone2.snaive<- snaive(yTrain2.ts, yValid2.ts, "Zone 2")
zone3.snaive<- snaive(yTrain3.ts, yValid3.ts, "Zone 3")
```

```{r}
zone1.snaive[[1]]
zone2.snaive[[1]]
zone3.snaive[[1]]
```

```{r}
ggarrange(zone1.snaive[[2]], zone2.snaive[[2]], zone3.snaive[[2]], ncol=1)
```


## Holt Winters


```{r}
#Built function for Holt Winters model
holtW<- function(yTrain.ts, yValid.ts, titl) {
  holt.model <- ets(yTrain.ts, model = "ZAA", alpha = .2, gamma = .05)
  holt.model.pred <- forecast(holt.model, h = nValid)
  holt.model.pred.ts <- ts(holt.model.pred$mean, start = c(45, 1), end= c(52, 7), frequency = 7)
  model <- holt.model
  p <- autoplot(yTrain.ts) + 
    autolayer( holt.model.pred, color = 'red')+
    autolayer(yValid.ts, color = "blue")+
    labs(title =titl,x = "week", y = "Power Consumption")+
    theme (title =element_text(size=7))
  acc <- accuracy(holt.model.pred, yValid.ts)
  lst <- list(acc,p, model)
  return(lst)}
#call model output
zone1.holt<- holtW(yTrain1.ts, yValid1.ts, "Zone 1")
zone2.holt<- holtW(yTrain2.ts, yValid2.ts, "Zone 2")
zone3.holt<- holtW(yTrain3.ts, yValid3.ts, "Zone 3")
```

```{r}
zone1.holt[[1]]
zone2.holt[[1]]
zone3.holt[[1]]
```

```{r}
ggarrange(zone1.holt[[2]], zone2.holt[[2]], zone3.holt[[2]], ncol=1)
```

```{r}
#Residuals plot for Holt model
plot(zone1.holt[[3]]$residuals, xlab = "Year", ylab = "Residuals", bty = "l", 
    lwd = 1, main = "Residuals of Zone 1 power consumption- Holt's Model")
```


## TSLM 

```{r}
#Built function for TSLM
tslm<- function(yTrain.ts, yValid.ts, titl) {
  #Create the formula for the regression model
  (formula <- as.formula(paste("yTrain1.ts", paste(c("trend", "season", 
                                                   colnames(xTrain)),
                                                 collapse = "+"), sep = "~")))
  tslm.model <- forecast::tslm(formula, data = xTrain)
  tslm.model.pred <- forecast(tslm.model, newdata = xValid)
  tslm.model.pred.ts <- ts(tslm.model.pred$mean, start = c(45, 1), end= c(52, 7), frequency = 7)
  p <- autoplot(yTrain.ts) + 
    autolayer( tslm.model.pred, color = 'red')+
    autolayer(yValid.ts, color = "blue")+
    labs(title =titl,x = "week", y = "Power Consumption")+
    theme (title =element_text(size=7))
  acc <- accuracy(tslm.model.pred, yValid.ts)
  lst <- list(acc,p)
  return(lst)}
#call model output
zone1.tslm<- tslm(yTrain1.ts, yValid1.ts, "Zone 1")
zone2.tslm<- tslm(yTrain2.ts, yValid2.ts, "Zone 2")
zone3.tslm<- tslm(yTrain3.ts, yValid3.ts, "Zone 3")
```

```{r}
zone1.tslm[[1]]
zone2.tslm[[1]]
zone3.tslm[[1]]
```

```{r}
ggarrange(zone1.tslm[[2]], zone2.tslm[[2]], zone3.tslm[[2]], ncol=1)
```

## Auto Arima Model

```{r}
#Built function for TSLM model in 3 zones
autoArima<- function(yTrain.ts, yValid.ts, titl) {
  predictor_train <- as.matrix(xTrain)
  predictors_test <- as.matrix(xValid)
  autoArima.model <- forecast::auto.arima(yTrain1.ts, xreg =predictor_train)
  autoArima.model.pred <- forecast(autoArima.model, h = nValid, xreg= predictors_test)
  autoArima.model.pred.ts <- ts(autoArima.model.pred$mean, start = c(45, 1), end= c(52,7), frequency = 7)
  model <- autoArima.model
  p <- autoplot(yTrain.ts) + 
    autolayer( autoArima.model.pred, color = 'red')+
    autolayer(yValid.ts, color = "blue")+
    labs(title =titl,x = "week", y = "Power Consumption")+
    theme (title =element_text(size=7))
  acc <- accuracy(autoArima.model.pred, yValid.ts)
  lst <- list(acc,p,model)
  return(lst)}
#call model output
zone1.autoArima<- autoArima(yTrain1.ts, yValid1.ts, "Zone 1")
zone2.autoArima<- autoArima(yTrain2.ts, yValid2.ts, "Zone 2")
zone3.autoArima<- autoArima(yTrain3.ts, yValid3.ts, "Zone 3")
```

```{r}
#Output auto arima performance
zone1.autoArima[[1]]
zone2.autoArima[[1]]
zone3.autoArima[[1]]
```

```{r}
ggarrange(zone1.autoArima[[2]], zone2.autoArima[[2]], zone3.autoArima[[2]], ncol=1)
```

```{r}
#Output auto arima parameters
zone1.autoArima[[3]]
zone2.autoArima[[3]]
zone3.autoArima[[3]]
```

### Optimized ARIMA model on Zone 1

```{r}
#Plot autocorrelation of yTrain1.ts
ggtsdisplay(yTrain1.ts)
```

```{r}
#Plot autocorrelation of differenced yTrain1.ts after two non-seasonal differences and 1 seasonal difference
yTrain1.ts %>% diff %>% diff%>% diff(lag = 7) %>% ggtsdisplay(lag=44)
```

After taking two non-seasonal differences and 1 seasonal difference,the autocorrelation seem improved and has removed some of the seasonality.  

```{r}
#Built function for my Arima models in 3 zones
myArima<- function(yTrain.ts, yValid.ts, titl) {
  predictor_train <- as.matrix(xTrain)
  predictors_test <- as.matrix(xValid)
  myArima.model <- forecast::Arima(yTrain1.ts,  order = c(1,2,1), 
                                   seasonal=c(1,0,0), xreg =predictor_train)
  myArima.model.pred <- forecast(myArima.model, h = nValid, xreg= predictors_test)
  myArima.model.pred.ts <- ts(myArima.model.pred$mean, start = c(45, 1), end= c(52,7), frequency = 7)
  model <- myArima.model
  p <- autoplot(yTrain.ts, series='Train', color = "black") + 
    autolayer( myArima.model.pred, color = 'red', series='Forecast')+
    autolayer(yValid.ts, color = "blue", series = 'Test')+
    labs(title =titl,x = "week", y = "Power Consumption")+
    theme (title =element_text(size=7))+
    guides(color = guide_legend(title = "Data Series")) +
    scale_color_manual(values = c(Train = "black", Forecast = "red", 
                                Test = "blue"))
    #coord_cartesian(xlim = c(15, 54))
  acc <- accuracy(myArima.model.pred, yValid.ts)
  lst <- list(acc,p, model)
  return(lst)}
#call model output
zone1.myArima<- myArima(yTrain1.ts, yValid1.ts, "Zone 1")
zone2.myArima<- myArima(yTrain2.ts, yValid2.ts, "Zone 2")
zone3.myArima<- myArima(yTrain3.ts, yValid3.ts, "Zone 3")
```

```{r}
#Output auto arima performance
zone1.myArima[[1]]
zone2.myArima[[1]]
zone3.myArima[[1]]
```

```{r}
grid.arrange(zone1.myArima[[2]], zone2.myArima[[2]], zone3.myArima[[2]], ncol=1)
```

### Residual plots for Holt's and Arima model

```{r}
par(mfrow=c(2,1))
#Residuals plot for Holt model
plot(zone1.holt[[3]]$residuals, xlab = "Year", ylab = "Residuals", bty = "l", 
    lwd = 1, main = "Residuals of Zone 1 power consumption- Holt's Model")
#Residuals plot for arima model
plot(zone1.myArima[[3]]$residuals, xlab = "Year", ylab = "Residuals", bty = "l", 
    lwd = 1, main = "Residuals of Zone 1 power consumption- Arima Model")
```

### Neuro Network

```{r}
#Built function for my Arima models in 3 zones
nnetar<- function(yTrain.ts, yValid.ts, titl) {
  nnetar.model <- forecast::nnetar(yTrain1.ts)
  nnetar.model.pred <- forecast(nnetar.model, h = nValid)
  #nnetar.model.pred.ts <- ts(nnetar.model.pred, start = c(45, 1), end= c(52,7), frequency = 7)
  p <- autoplot(yTrain.ts, series='Train', color = "black") + 
    autolayer( nnetar.model.pred, color = 'red', series='Forecast')+
    autolayer(yValid.ts, color = "blue", series = 'Test')+
    labs(title =titl,x = "week", y = "Power Consumption")+
    theme (title =element_text(size=7))+
    guides(color = guide_legend(title = "Data Series")) +
    scale_color_manual(values = c(Train = "black", Forecast = "red", 
                                Test = "blue"))
    #coord_cartesian(xlim = c(15, 54))
  acc <- accuracy(nnetar.model.pred, yValid.ts)
  lst <- list(acc,p)
  return(lst)}
#call model output
zone1.nnetar<- nnetar(yTrain1.ts, yValid1.ts, "Zone 1")
zone2.nnetar<- nnetar(yTrain2.ts, yValid2.ts, "Zone 2")
zone3.nnetar<- nnetar(yTrain3.ts, yValid3.ts, "Zone 3")
```

```{r}
#Output auto arima performance
zone1.nnetar[[1]]
zone2.nnetar[[1]]
zone3.nnetar[[1]]
```

```{r}
grid.arrange(zone1.nnetar[[2]], zone2.myArima[[2]], zone3.nnetar[[2]], ncol=1)
```

## Provide two weeks future forecast

```{r}
fc.period = 14
#Convert full data to ts
y1.ts <- ts(y1, start = c(1, 1), end= c(52, 7), frequency = 7)
full.holt.model <- ets(y1.ts, model = "ZAA", alpha = .2, gamma = .05)
full.holt.model.pred <- forecast(full.holt.model, h = fc.period)
full.holt.model.pred.ts <- ts(full.holt.model.pred$mean, start = c(45, 1), end= c(52, 7), frequency = 7)
    autoplot(y1.ts) + 
    autolayer( full.holt.model.pred, color = 'orange')+
    labs(title = "Two week Forecast Plot",x = "week", y = "Power Consumption")+
    theme (title =element_text(size=7))
```

```{r}
#Forecast Value
full.holt.model.pred$mean
```

```{r}

```



